---
title: "Coursework 1: MA22019, Introduction to Data Science"
author: "Replace With Your Name"
format: gfm
---

<!--- DO NOT DELETE THIS LINE --->
<!--- METADATA ANCHOR: MA22019-CW1-2026 --->

Please read the `coursework_01_instructions.pdf` for the full context, data dictionary, and detailed task requirements before starting.



```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(tidytext)
library(knitr)
```

## Part A: Data Wrangling & EDA 

### Task 1: Data Joining
<!--- DO NOT DELETE THIS LINE - TASK 1 ANCHOR --->

```{r}
#| label: task-1-code
# Write your Task 1 code here:

history <- read_csv("data/history.csv")
titles <- read_csv("data/titles.csv")
users <- read_csv("data/users.csv")

history_clean <- history %>% unique ()
titles_clean <- titles %>% unique()
users_clean <- users %>% unique()

cinemetrics <- history_clean %>% 
                  left_join (users_clean, by = 'user_id') %>% 
                    left_join(titles_clean , by = 'movie_id')

nrow(history)
nrow(history_clean)
nrow(titles)
nrow(titles_clean)
nrow(cinemetrics)

```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It was necessary to remove duplicates from history to avoid counting the same viewing record more than once, and duplicates had to be removed from the titles data frame because otherwise repeated movie_id rows would create duplicate matches in the join and artificially increase the final row count.


### Task 2: Missing Values Handling
<!--- DO NOT DELETE THIS LINE - TASK 2 ANCHOR --->

```{r}
#| label: task-2-code
# Write your Task 2 code here:

missing_summary <- tibble (
  variable = colnames ( cinemetrics) , 
  missing_values = colSums(is.na(cinemetrics))
) %>%  arrange(desc(missing_values))

missing_summary %>% knitr::kable(caption = "Missing values per variable")

user_rating_median <- median(cinemetrics$user_rating_100 , na.rm = TRUE)
user_rating_median

```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It is important to use explicit missing-value functions because they directly identify and count NA values, whereas generic overview functions may not clearly report missingness, which can lead to misinterpretation of the dataset and misleading summary statistics such as the mean, median, or mode; this issue is especially common for character variables, where missing values are often not clearly flagged.This happens because those columns are treated as text rather than numerical variables, so the output may only preview values instead of explicitly counting NAs.


### Task 3: Genre Grouping
<!--- DO NOT DELETE THIS LINE - TASK 3 ANCHOR --->

```{r}
#| label: task-3-code
# Write your Task 3 code here:

cinemetrics |>
  count(genre_tags, sort = TRUE)

cinemetrics <- cinemetrics %>% 
  mutate ( primary_genre = genre_tags %>% 
             as_factor() %>% 
             fct_collapse(
               'Animation' = c("animated", "animation", "animation, kids"),
               'Comedy' = c("comedy", "comedy, family"),
              'Sci-Fi' = c("Science fiction", "sci-fi", "sci-fi, fantasy"),
              'Drama' = c("Drama", "drama, emotional", "drama, critically-acclaimed"),
              'Horror' = c("horror", "scary", "terror"),
              'Romance' = c("Romance", "rom-com", "romantic", "romantic comedy"),
              'Action' = c("action", "action, high-octane"),
              'Documentary' = c("Documentary", "documentary, real-world")
              )
  )

cinemetrics <-  cinemetrics %>%
  mutate( primary_genre = fct_lump(primary_genre , n = 5 , other_level = 'other'))

levels(cinemetrics$primary_genre)
cinemetrics %>% count(primary_genre, sort = TRUE)



```

```{r}
#| label: task-3-verification
# Write your verification code here:

rating_verification <- cinemetrics %>% 
  filter ( genre_tags == 'Romance' | genre_tags =='romantic comedy') %>% 
  group_by( genre_tags) %>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE),
              n = n())

rating_verification %>%
  knitr::kable(caption = "Comparison of raw Romance and romantic comedy genre tags")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

Collapsing the romantic comedy subgroup into the broader Romance category masks its lower average rating, because its mean score of 66.35 across 145 observations is diluted within the much larger Romance group, which has a higher average rating of 69.83 across 991 observations. This is called aggregation distortion.


### Task 4: Genre Ratings Table
<!--- DO NOT DELETE THIS LINE - TASK 4 ANCHOR --->

```{r}
#| label: task-4-code
# Write your Task 4 code here:

genre_ratings <- cinemetrics %>% 
  group_by(primary_genre)%>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE)) %>%   arrange(desc (avg_user_rating_100))

genre_ratings %>% knitr::kable ( caption = "Average User rating for Primary genre")


```

```{r}
#| label: task-4-verification
# Write your verification code here:

horror_budget <- cinemetrics %>% 
    filter ( genre_tags == "horror" | genre_tags == "scary" |genre_tags== "terror" ) %>% 
    summarise( genre = 'Horror' , avg_budget = mean(budget_usd, na.rm= TRUE)
                 )

action_budget <- cinemetrics %>% 
  filter ( genre_tags == "action" | genre_tags == "action, high-octane" ) %>% 
  summarise( genre = "Action" , avg_budget = mean(budget_usd , na.rm = TRUE ))

SciFi_budget <- cinemetrics %>% 
  filter ( primary_genre == "Sci-Fi") %>% 
  summarise( genre = 'Sci-Fi' , avg_budget = mean (budget_usd, na.rm = TRUE) )

avg_budget_usd = bind_rows (horror_budget,action_budget,SciFi_budget ) %>% 
  arrange ( desc(avg_budget))

avg_budget_usd %>%  knitr::kable (caption = " Average production budget (USD)")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The empirical results show that a higher production budget does not guarantee a higher average user rating, because Sci-Fi has the highest average budget of 139588009 USD yet does not have the highest average rating (69.69) in the genre ratings table, where Other performs better(74.34) . However, Action and Horror are absorbed into Other, so their separate average ratings cannot be directly observed from the collapsed table. This demonstrates how a massive generalised category can hide subgroup nuances. The poor performance of a small niche film group, may be concealed when it is absorbed into a larger category whose overall mean is driven by other films.


### Task 5: Ratings and Engagement Scores
<!--- DO NOT DELETE THIS LINE - TASK 5 ANCHOR --->

```{r}
#| label: task-5-code
# Write your Task 5 code here:

top_10_movies <- cinemetrics %>% 
  count(movie_title , sort = TRUE) %>% 
  top_n(10)

top_10_ratings <- cinemetrics %>% 
  inner_join(top_10_movies, by = "movie_title" ) %>% 
  mutate ( movie_title = fct_reorder(movie_title,user_rating_100,.fun = median, .desc = TRUE,.na_rm = TRUE))
  
ggplot (data = top_10_ratings , aes (x = "", y = user_rating_100)) + geom_boxplot() + facet_wrap( ~movie_title , nrow = 2)
```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

Up contains the lowest individual user rating among the 10 most-watched movies, as shown by the lowest outlier point on the plot which falls below 25. Its relatively narrow box also suggests a smaller interquartile range approximately from 58 to 72, meaning the middle 50% of ratings are more tightly clustered and audience opinion is more consistent.

```{r}
#| label: task-5-verification
# Write your verification code here:

top_10_ratings_IQR <- top_10_ratings %>% 
  group_by(movie_title) %>% 
  summarize( IQR_rating = IQR(user_rating_100,na.rm = TRUE) ) %>% 
  arrange(desc(IQR_rating))
  
top_10_ratings_IQR %>% knitr::kable(caption = "IQR of user ratings for 10 most watched movies")


```

#### Empirical Verification

Inside Out has the highest IQR range of 22.5 and Dune has the lowest IQR range of 18.


### Task 6: Engagement Score and Subscription Type
<!--- DO NOT DELETE THIS LINE - TASK 6 ANCHOR --->

```{r}
#| label: task-6-code
# Write your Task 6 code here:

engagement_2d <- cinemetrics %>%  
  filter (!is.na(subscription_type)) %>% 
  group_by(subscription_type) %>% 
  summarise(avg_engagement_score = mean(engagement_score, na.rm = TRUE))

engagement_2d %>% knitr::kable(caption = "Average engagement score by subscription type")

age_quartiles <- cinemetrics$age %>% 
  quantile(probs = c(0.25,0.50,0.75), na.rm = TRUE )

age_quartiles

cinemetrics <- cinemetrics %>% 
  mutate( age_tier = case_when(age <= 25 ~ "Q1",
                               age <= 35 ~ "Q2",
                               age <= 45 ~ "Q3", 
                               age > 45 ~ "Q4" ))

engagement_3d <- cinemetrics %>% 
  filter (!is.na(subscription_type), !is.na(age_tier)) %>% 
  group_by(subscription_type , age_tier) %>% 
  summarise(avg_engagement_score = mean(engagement_score,na.rm = TRUE))


```

```{r}
#| label: task-6-verification
# Write your verification code here:

engagement_verification <- cinemetrics %>% 
  filter (!is.na(subscription_type), !is.na(age_tier)) %>% 
  group_by(subscription_type , age_tier) %>% 
  summarise(user_count = n ())

engagement_verification %>% 
tidyr :: pivot_wider( names_from = subscription_type , values_from = user_count)


```

#### Empirical Verification

Basic users are concentrated in the younger age groups, with 4958 users in Q1 (under 25 years old) and 3523 in Q2 (25 to 35 years old), while Premium users are concentrated in the older age groups, with 3762 users in Q3 (35 to 45 years old) and 4178 in Q4 (over 45 years old). This shows that the two subscription types are distributed unevenly across age, which is likely to distort the overall comparison of engagement scores.

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The 2D analysis suggests that premium users are engaged more overall with an average engagement score of 56.48 whereas basic users have an average engagement score of 44.04 overall.However, the 3D analysis shows the opposite pattern within every age tier, where Basic users have higher average engagement than Premium users in Q1 (28.94 vs 12.66), Q2 (50.93 vs 38.24), Q3 (74.34 vs 55.10), and Q4 (80.10 vs 68.77). This is called the Simpsons paradox where an overall trend reverses after controlling for a third variable. It occurs here because the subscription types are distributed very unevenly across age tiers: Basic users are concentrated in the younger, lower-engagement groups, while Premium users are concentrated in the older, higher-engagement groups. As a result, the overall engagement averages are driven by differences in age composition rather than subscription type alone.


---

## Part B: Text Data Analysis

### Task 7: Tokenization & Word Frequency
<!--- DO NOT DELETE THIS LINE - TASK 7 ANCHOR --->

```{r}
#| label: task-7-code
# Write your Task 7 code here:

tidy_review = cinemetrics %>% 
  filter(!is.na(review_snippet)) %>%
  select(review_snippet) %>% 
  unnest_tokens(output = word , input = review_snippet) %>% 
  anti_join (stop_words , by ="word") %>% 
  count (word, sort = TRUE )

top_10_words = tidy_review %>% 
  head(10)

top_10_words

ggplot(top_10_words, aes( x= reorder(word,n) , y = n )) + geom_col() + labs( x = "Word" , y = "Frequency" , title = " Frequency of Top 10 Most Occuring Words") +coord_flip()



```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The two most frequent non-stop words are film and watch with a frequency of 5674 and 5317 respectively, but these words provide very little useful insight because they are generic terms that are naturally expected to appear often in movie reviews. They do not explain why audiences rated a movie highly or what makes a specific genre distinctive, since they are not tied to particular themes, emotions, or stylistic features. As a result, simple word frequency is limited here, and a measure such as TF-IDF is more useful for identifying words that are more distinctive and informative within particular genres or review groups.


### Task 8: Sentiment Analysis
<!--- DO NOT DELETE THIS LINE - TASK 8 ANCHOR --->

```{r}
#| label: task-8-code
# Write your Task 8 code here:

bing_lexicon <- get_sentiments("bing")

review_sentiment <-  cinemetrics %>% 
  filter(!is.na(review_snippet),!is.na(primary_genre),!is.na(user_rating_100)) %>%
  mutate(review_id = row_number()) %>% 
  select(review_id,primary_genre,user_rating_100,review_snippet) %>% 
  unnest_tokens(output = word , input = review_snippet) %>% 
  inner_join(bing_lexicon, by= "word") %>% 
  mutate( sentiment_value = if_else(sentiment=="positive", +1 , -1 )) %>% 
  group_by(review_id,primary_genre,user_rating_100) %>% 
  summarise(net_sentiment= sum(sentiment_value))

genre_sentiment <- review_sentiment %>% 
  group_by(primary_genre) %>% 
  summarise( avg_net_sentiment = mean(net_sentiment) , 
             avg_user_rating = mean (user_rating_100))

ggplot(genre_sentiment, aes(x=avg_net_sentiment, y = avg_user_rating ,colour = primary_genre )) + geom_point()


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The clear anomaly on the scatter plot is other, which sits far to the left with a negative average net sentiment of roughly -0.2, yet has the highest average user rating at around 74.5. This differs from the rest of the genres, which cluster together with positive average sentiment values around 0.75 to 0.80 and ratings near 71 to 71.5.

This semantic anomaly occurs because the Bing lexicon classifies words individually as positive or negative without considering context, so reviews in other may contain lexically negative words that are used in a praising or genre-specific way. 

In addition, because other is a collapsed category containing multiple smaller genres, its language patterns are likely to be more mixed and less well captured by a simple sentiment lexicon.

For example, words such as “terrifying” or “disturbing” may be labelled as negative, even though in the context of a horror or intense film review they can actually express praise. As a result, the average sentiment score may appear artificially low even when the films are highly rated by audiences.


### Task 9: Advanced Text Mining (TF-IDF)
<!--- DO NOT DELETE THIS LINE - TASK 9 ANCHOR --->

```{r}
#| label: task-9-code
# Write your Task 9 code here: 


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

1. *Replace this text with the TF-IDF score for the word "film".*
2. *Replace this text with your mathematical explanation using the IDF equation from the lecture.*
3. *Replace this text with your comparison of TF-IDF vs frequency.*



### Task 10: Executive Recommendation
<!--- DO NOT DELETE THIS LINE - TASK 10 ANCHOR --->

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your one-paragraph executive recommendation (max 150 words), citing at least three specific findings from earlier tasks and acknowledging one limitation.*


---
<!--- WORD COUNT ANCHOR --->
```{r}
#| label: word-count
#| echo: false
#| message: false
#| warning: false

count_prose_words <- function(file_path) {
  # Read the raw lines of the file
  lines <- readLines(file_path, warn = FALSE)
  
  # 1. Strip YAML Header (Lines between the first two '---')
  yaml_bounds <- grep("^---$", lines)
  if (length(yaml_bounds) >= 2) {
    lines <- c(lines[1:(yaml_bounds[1]-1)], lines[(yaml_bounds[2]+1):length(lines)])
  }
  
  # 2. Strip Code Chunks (Lines between pairs of '```')
  chunk_bounds <- grep("^```", lines)
  if (length(chunk_bounds) %% 2 == 0 && length(chunk_bounds) > 0) {
    # Create a vector of line indices to remove
    lines_to_remove <- integer(0)
    for (i in seq(1, length(chunk_bounds), by = 2)) {
      lines_to_remove <- c(lines_to_remove, chunk_bounds[i]:chunk_bounds[i+1])
    }
    lines <- lines[-lines_to_remove]
  }
  
  # 3. Strip Markdown Headers (Lines starting with #)
  lines <- lines[!grepl("^#+", trimws(lines))]
  
  # 4. Strip HTML Comments (Lines starting with <!--)
  lines <- lines[!grepl("^<!--", trimws(lines))]
  
  # 5. Combine what's left into a single string
  text_only <- paste(lines, collapse = " ")
  
  # 6. Remove common markdown characters that might be counted as words
  text_only <- gsub("[*>`-]", " ", text_only)
  
  # 7. Count the remaining words
  words <- unlist(strsplit(trimws(text_only), "\\s+"))
  words <- words[words != ""]
  
  return(length(words))
}

# Run the function on the current file
true_word_count <- count_prose_words(knitr::current_input())

# Print the final result formatted nicely
cat(sprintf("**Prose Word Count:** %d words", true_word_count))
```
