---
title: "Coursework 1: MA22019, Introduction to Data Science"
author: "Replace With Your Name"
format: gfm
---

<!--- DO NOT DELETE THIS LINE --->
<!--- METADATA ANCHOR: MA22019-CW1-2026 --->

Please read the `coursework_01_instructions.pdf` for the full context, data dictionary, and detailed task requirements before starting.



```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(tidytext)
library(knitr)
```

## Part A: Data Wrangling & EDA 

### Task 1: Data Joining
<!--- DO NOT DELETE THIS LINE - TASK 1 ANCHOR --->

```{r}
#| warning: false
#| message: false
#| label: task-1-code
# Write your Task 1 code here:

history <- read_csv("data/history.csv")
titles <- read_csv("data/titles.csv")
users <- read_csv("data/users.csv")

history_clean <- history %>% unique ()
titles_clean <- titles %>% unique()
users_clean <- users %>% unique()

cinemetrics <- history_clean %>% 
                  left_join (users_clean, by = 'user_id') %>% 
                    left_join(titles_clean , by = 'movie_id')

nrow(history)
nrow(history_clean)
nrow(titles)
nrow(titles_clean)
nrow(cinemetrics)



```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It was necessary to remove duplicates from history to avoid counting the same viewing record more than once, and duplicates had to be removed from the titles data frame because otherwise repeated movie_id rows would create duplicate matches in the join and artificially increase the final row count.


### Task 2: Missing Values Handling
<!--- DO NOT DELETE THIS LINE - TASK 2 ANCHOR --->

```{r}
#| label: task-2-code
# Write your Task 2 code here:

missing_summary <- tibble (
  variable = colnames ( cinemetrics) , 
  missing_values = colSums(is.na(cinemetrics))
) %>%  arrange(desc(missing_values))

missing_summary %>% knitr::kable(caption = "Missing values per variable")

user_rating_median <- median(cinemetrics$user_rating_100 , na.rm = TRUE)
user_rating_median

```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It is mathematically necessary to use explicit missing-value functions because they directly test whether each observation is missing and allow the exact number of NA values to be counted. Generic data overview functions provide a broad summary of the dataset, so missingness may be overlooked and summary statistics may be misinterpreted. This problem is especially common for character variables, where missing values are often not clearly flagged because the column is treated as text rather than being numerically summarised.


### Task 3: Genre Grouping
<!--- DO NOT DELETE THIS LINE - TASK 3 ANCHOR --->

```{r}
#| label: task-3-code
# Write your Task 3 code here:

cinemetrics %>% 
  count(genre_tags, sort = TRUE)

cinemetrics <- cinemetrics %>% 
  mutate ( primary_genre = genre_tags %>% 
             as_factor() %>% 
             fct_collapse(
               'Animation' = c("animated", "animation", "animation, kids"),
               'Comedy' = c("comedy", "comedy, family"),
              'Sci-Fi' = c("Science fiction", "sci-fi", "sci-fi, fantasy"),
              'Drama' = c("Drama", "drama, emotional", "drama, critically-acclaimed"),
              'Horror' = c("horror", "scary", "terror"),
              'Romance' = c("Romance", "rom-com", "romantic", "romantic comedy"),
              'Action' = c("action", "action, high-octane"),
              'Documentary' = c("Documentary", "documentary, real-world")
              )
  )


levels(cinemetrics$primary_genre)
cinemetrics %>% count(primary_genre, sort = TRUE)



```
##### comment: No additional sub-categories remained, so an "Other" category was not needed and the data contained exactly 8 levels.

```{r}
#| label: task-3-verification
# Write your verification code here:

rating_verification <- cinemetrics %>% 
  filter ( genre_tags == 'Romance' | genre_tags =='romantic comedy') %>% 
  group_by( genre_tags) %>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE),
              n = n())

rating_verification %>%
  knitr::kable(caption = "Comparison of raw Romance and Romantic Comedy genre tags")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

Collapsing the romantic comedy subgroup into the broader Romance category masks its lower average rating, because its mean score of 66.35 across 145 observations is diluted within the much larger Romance group, which has a higher average rating of 69.83 across 991 observations. This is called aggregation distortion.


### Task 4: Genre Ratings Table
<!--- DO NOT DELETE THIS LINE - TASK 4 ANCHOR --->

```{r}
#| label: task-4-code
# Write your Task 4 code here:

genre_ratings <- cinemetrics %>% 
  group_by(primary_genre)%>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE)) %>%   arrange(desc (avg_user_rating_100))

genre_ratings %>% knitr::kable ( caption = "Average User rating for Primary genre")


```

```{r}
#| label: task-4-verification
# Write your verification code here:


budget_verification <- cinemetrics %>% 
  filter(primary_genre=="Action"| primary_genre =="Sci-Fi"|primary_genre == "Horror") %>% 
  group_by(primary_genre) %>% 
  summarise( avg_budget_usd = mean(budget_usd, na.rm= TRUE), .groups = "drop")%>% 
  arrange(desc(avg_budget_usd))
  
budget_verification %>% knitr::kable(caption = "Average Budget in USD")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The empirical results show that a higher production budget does not guarantee a higher average user rating. Sci-Fi has the highest average budget at 139,588,009 USD, and an average user rating of 69.69 however Horror has by far the lowest average budget at 10,181,094 USD but the highest average user rating at 79.78. This suggests that higher spending alone does not determine average user rating.

A broad category such as Action may hide important sub-genre differences because averaging across many films can conceal weaker niche groups within the larger category. This is similar to the pattern seen earlier in Task 3, where Romantic Comedy (66.35) was masked within the broader Romance category (69.83). More generally, the poor performance of a small niche film group can be hidden when it is absorbed into a large collapsed category whose average is driven by many other films.

### Task 5: Ratings and Engagement Scores
<!--- DO NOT DELETE THIS LINE - TASK 5 ANCHOR --->

```{r}
#| warning: false
#| message: false
#| label: task-5-code
# Write your Task 5 code here:


top_10_movies <- cinemetrics %>% 
  count(movie_title , sort = TRUE) %>% 
  top_n(10)

top_10_ratings <- cinemetrics %>% 
  inner_join(top_10_movies, by = "movie_title" ) %>% 
  mutate ( movie_title = fct_reorder(movie_title,user_rating_100,.fun = median, .desc = TRUE,.na_rm = TRUE))
  
ggplot (data = top_10_ratings , aes (x = "", y = user_rating_100)) + geom_boxplot() + facet_wrap( ~movie_title , nrow = 1) + theme(
    strip.text = element_text(size = 5))


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

Up contains the lowest individual user rating among the 10 most-watched movies, as shown by the lowest outlier point on the plot which falls below 25. Its relatively narrow box also suggests a smaller interquartile range approximately from 58 to 72, meaning the middle 50% of ratings are more tightly clustered and audience opinion is more consistent.

```{r}
#| label: task-5-verification
# Write your verification code here:

top_10_ratings_IQR <- top_10_ratings %>% 
  group_by(movie_title) %>% 
  summarize( IQR_rating = IQR(user_rating_100,na.rm = TRUE) ) %>% 
  arrange(desc(IQR_rating))
  
top_10_ratings_IQR %>% knitr::kable(caption = "IQR of user ratings for 10 most watched movies")


```

#### Empirical Verification

Inside Out has the highest IQR range of 22.5 and Dune has the lowest of 18.


### Task 6: Engagement Score and Subscription Type
<!--- DO NOT DELETE THIS LINE - TASK 6 ANCHOR --->

```{r}
#| warning: false
#| message: false
#| label: task-6-code
# Write your Task 6 code here:

engagement_2d <- cinemetrics %>%  
  filter (!is.na(subscription_type)) %>% 
  group_by(subscription_type) %>% 
  summarise(avg_engagement_score = mean(engagement_score, na.rm = TRUE))

engagement_2d %>% knitr::kable(caption = "Average engagement score by subscription type")

age_quartiles <- cinemetrics$age %>% 
  quantile(probs = c(0.25,0.50,0.75), na.rm = TRUE )

age_quartiles

cinemetrics <- cinemetrics %>% 
  mutate( age_tier = case_when(age <= 25 ~ "Q1",
                               age <= 35 ~ "Q2",
                               age <= 45 ~ "Q3", 
                               age > 45 ~ "Q4" ))

engagement_3d <- cinemetrics %>% 
  filter (!is.na(subscription_type), !is.na(age_tier)) %>% 
  group_by(subscription_type , age_tier) %>% 
  summarise(avg_engagement_score = mean(engagement_score,na.rm = TRUE))

engagement_3d %>% knitr::kable(caption = "Average engagement score by subscription type and age quartiles")


```

```{r}
#| warning: false
#| message: false
#| label: task-6-verification
# Write your verification code here:

engagement_verification <- cinemetrics %>% 
  filter (!is.na(subscription_type), !is.na(age_tier)) %>% 
  group_by(subscription_type , age_tier) %>% 
  summarise(user_count = n ())

engagement_verification %>% 
tidyr :: pivot_wider( names_from = subscription_type , values_from = user_count)


```

#### Empirical Verification

Basic users are concentrated in the younger age groups, with 4958 users in Q1 (under 25 years old) and 3523 in Q2 (25 to 35 years old), while Premium users are concentrated in the older age groups, with 3762 users in Q3 (35 to 45 years old) and 4178 in Q4 (over 45 years old). This shows that the two subscription types are distributed unevenly across age, which is likely to distort the overall comparison of engagement scores.

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The 2D analysis suggests that Premium users are more engaged overall, with an average engagement score of 56.48 compared with 44.04 for Basic users. However, the 3D analysis shows the opposite pattern within every age tier: Basic users have higher average engagement than Premium users in Q1 (28.94 vs 12.66), Q2 (50.93 vs 38.24), Q3 (74.34 vs 55.10), and Q4 (80.10 vs 68.77). This is an example of Simpson’s paradox, where an overall trend reverses after controlling for a third variable. It occurs here because subscription types are distributed unevenly across age tiers (confounding variable) : Basic users are concentrated in younger, lower-engagement groups, while Premium users are concentrated in older, higher-engagement groups. Hence, the overall averages reflect differences in age composition rather than the effect of subscription type alone.


---

## Part B: Text Data Analysis

### Task 7: Tokenization & Word Frequency
<!--- DO NOT DELETE THIS LINE - TASK 7 ANCHOR --->

```{r}
#| label: task-7-code
# Write your Task 7 code here:

tidy_tokenized_review <- cinemetrics %>% 
  filter(!is.na(review_snippet)) %>%
  select(review_snippet) %>% 
  unnest_tokens(output = word , input = review_snippet) %>% 
  anti_join (stop_words , by ="word") %>% 
  count (word, sort = TRUE )

top_10_words <- tidy_tokenized_review %>% 
  head(10)

top_10_words

ggplot(top_10_words, aes( x= reorder(word,n) , y = n )) + geom_col() + labs( x = "Word" , y = "Frequency" , title = " Frequency of Top 10 Most Occuring Words") +coord_flip()



```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The two most frequent non-stop words are “film” and “watch”, with frequencies of 5674 and 5317 respectively. However, these words provide little insight because they are generic terms that appear often in movie reviews. They don't explain what makes a genre distinctive, since they are not linked to specific themes or emotions. This shows that simple word frequency is limited in this context, whereas TF-IDF is more useful for identifying words that are more distinctive within particular genres.


### Task 8: Sentiment Analysis
<!--- DO NOT DELETE THIS LINE - TASK 8 ANCHOR --->

```{r}
#| warning: false
#| message: false
#| label: task-8-code
# Write your Task 8 code here:

bing_lexicon <- get_sentiments("bing")

review_sentiment <-  cinemetrics %>% 
  filter(!is.na(review_snippet),!is.na(primary_genre),!is.na(user_rating_100)) %>%
  mutate(review_id = row_number()) %>% 
  select(review_id,primary_genre,user_rating_100,review_snippet) %>% 
  unnest_tokens(output = word , input = review_snippet) %>% 
  inner_join(bing_lexicon, by= "word") %>% 
  mutate( sentiment_value = if_else(sentiment=="positive", +1 , -1 )) %>% 
  group_by(review_id,primary_genre,user_rating_100) %>% 
  summarise(net_sentiment= sum(sentiment_value))

genre_sentiment <- review_sentiment %>% 
  group_by(primary_genre) %>% 
  summarise( avg_net_sentiment = mean(net_sentiment) , 
             avg_user_rating = mean (user_rating_100))

ggplot(genre_sentiment, aes(x=avg_net_sentiment, y = avg_user_rating ,colour = primary_genre )) + geom_point()


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The clear anomaly on the scatter plot is Horror, which has a negative average net sentiment of about -1.12 but the highest average user rating at roughly 78.5. This contrasts with the other genres, which cluster around positive sentiment values of 0.75 to 0.80 and ratings near 71 to 71.5.

This occurs because the Bing lexicon classifies words as positive or negative without considering context. In Horror reviews, words like “terrifying” or “disturbing” are labelled as negative even when they are praise.

In addition, Horror is a collapsed category containing several sub-genres, so its language is varied and less accurately captured by a simple sentiment lexicon. Hence, its average sentiment score may appear artificially low despite strong audience ratings.


### Task 9: Advanced Text Mining (TF-IDF)
<!--- DO NOT DELETE THIS LINE - TASK 9 ANCHOR --->

```{r}
#| label: task-9-code
# Write your Task 9 code here: 


# method 1 (works fyn with n but gives non zero tfidf for all romance and primary_genres and non zeroes only for others )

unique_words <- cinemetrics %>%
  filter(!is.na(review_snippet), !is.na(primary_genre)) %>%
  unnest_tokens(output = word, input = review_snippet) %>%
  anti_join(stop_words, by = "word") %>%
  count(primary_genre, word, sort = TRUE) %>% 
  bind_tf_idf(term = word , document = primary_genre , n = n ) %>% 
  arrange(primary_genre, desc(tf_idf))
  
top_unique_word <- unique_words %>%
  group_by(primary_genre) %>%
  arrange(desc(tf_idf), .by_group = TRUE) %>%
  slice_head(n = 1) %>%
  select(primary_genre, word, tf_idf)
  
romance_unique <-  unique_words %>%
  filter(primary_genre == "Romance")

top_unique_word %>% knitr::kable(caption = "Highest TF-IDF word for each genre")

romance_unique_top <- romance_unique %>% 
  arrange(desc(tf_idf)) %>% 
  slice_head(n=3) %>% 
  select (primary_genre , word , tf_idf)

romance_unique_top %>% knitr :: kable (caption = "Top 3 TF-IDF words for Romance")

romance_common_top <- romance_unique %>% 
  arrange(desc(n)) %>% 
  slice_head(n=3) %>% 
  select(primary_genre , word , n)

romance_common_top %>% 
  knitr::kable(caption = "TOP 3 frequency words for Romance")
  

### for more better analysis lets try removing the word "watch" to see if there are any genre distinguishing words ###

unique_words_2<- cinemetrics %>%
  filter(!is.na(review_snippet), !is.na(primary_genre)) %>%
  unnest_tokens(output = word, input = review_snippet) %>%
  anti_join(stop_words, by = "word") %>%
  filter( word!= "watch") %>% 
  count(primary_genre, word, sort = TRUE) %>% 
  bind_tf_idf(term = word , document = primary_genre , n = n ) %>% 
  arrange(primary_genre, desc(tf_idf))

  
top_unique_word_2<- unique_words_2 %>%
  group_by(primary_genre) %>%
  arrange(desc(tf_idf), .by_group = TRUE) %>%
  slice_head(n = 1) %>%
  select(primary_genre, word, tf_idf)

romance_unique_2 <-  unique_words_2 %>%
  filter(primary_genre == "Romance")
```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

1. TF-IDF score for "film" is 0

2. This occurs because the inverse document frequency part of the TF-IDF formula is 0 for film. Although film appears frequently in Romance reviews, it also appears across all genre documents in the dataset, so its IDF becomes log(8/8) = log(1) = 0, which forces the overall TF-IDF score to 0.

3. The top 3 TF-IDF words in Romance are "watch", "average" and "fairly", with scores of 0.014, 0.0077 and 0.0077 respectively, whereas the top 3 frequency words are "film", "watch" and "average", with frequencies of 697, 680 and 359 respectively.

TF-IDF gives a better summary because it downweights words such as "film" that are common across all genres and therefore not distinctive to Romance. Although "watch" and "average" still appear in both lists, TF-IDF improves the summary by prioritising uniqueness as well as frequency.

As an additional exploratory check, I removed "watch", since it had the highest TF-IDF score in 7 of the 8 genres, but this produced limited change. Romance remained dominated by generic words such as "average", "fairly", and "bad".This suggests that Romance reviews contain generic vocabulary, so the difference is modest here. 

However, in genres like Horror the contrast is clearer, where a word like "scary" has a higher TF-IDF score (0.28) and provides a distinctive summary of the genre.


### Task 10: Executive Recommendation
<!--- DO NOT DELETE THIS LINE - TASK 10 ANCHOR --->

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

CineMetrics should focus on converting younger Basic users into Premium through targeted content and upgrade offers. Task 6 shows that although Premium users appear more engaged overall (56.48 vs 44.04), Basic users are more engaged within every age tier, while younger users are heavily concentrated in Basic subscriptions (Q1: 4958 vs 447; Q2: 3523 vs 1496), suggesting strong upgrade potential.

Task 4 also shows that Horror achieves the highest average rating (79.78) despite the lowest average budget (10,181,094 USD), indicating cost-effective investment potential.

Task 9 further suggests that Horror (Scary – Highest TF-IDF)  generates more distinctive audience language than broader genres (Watch – Highest TF-IDF) supporting its strategic value. 

One limitation is that short review snippets and Bing lexicon analysis may misrepresent genre-specific sentiment, since words such as “terrifying” may be classified as negative even when they reflect positive Horror reception



---
<!--- WORD COUNT ANCHOR --->
```{r}
#| label: word-count
#| echo: false
#| message: false
#| warning: false

count_prose_words <- function(file_path) {
  # Read the raw lines of the file
  lines <- readLines(file_path, warn = FALSE)
  
  # 1. Strip YAML Header (Lines between the first two '---')
  yaml_bounds <- grep("^---$", lines)
  if (length(yaml_bounds) >= 2) {
    lines <- c(lines[1:(yaml_bounds[1]-1)], lines[(yaml_bounds[2]+1):length(lines)])
  }
  
  # 2. Strip Code Chunks (Lines between pairs of '```')
  chunk_bounds <- grep("^```", lines)
  if (length(chunk_bounds) %% 2 == 0 && length(chunk_bounds) > 0) {
    # Create a vector of line indices to remove
    lines_to_remove <- integer(0)
    for (i in seq(1, length(chunk_bounds), by = 2)) {
      lines_to_remove <- c(lines_to_remove, chunk_bounds[i]:chunk_bounds[i+1])
    }
    lines <- lines[-lines_to_remove]
  }
  
  # 3. Strip Markdown Headers (Lines starting with #)
  lines <- lines[!grepl("^#+", trimws(lines))]
  
  # 4. Strip HTML Comments (Lines starting with <!--)
  lines <- lines[!grepl("^<!--", trimws(lines))]
  
  # 5. Combine what's left into a single string
  text_only <- paste(lines, collapse = " ")
  
  # 6. Remove common markdown characters that might be counted as words
  text_only <- gsub("[*>`-]", " ", text_only)
  
  # 7. Count the remaining words
  words <- unlist(strsplit(trimws(text_only), "\\s+"))
  words <- words[words != ""]
  
  return(length(words))
}

# Run the function on the current file
true_word_count <- count_prose_words(knitr::current_input())

# Print the final result formatted nicely
cat(sprintf("**Prose Word Count:** %d words", true_word_count))
```
