---
title: "Coursework 1: MA22019, Introduction to Data Science"
author: "Replace With Your Name"
format: gfm
---

<!--- DO NOT DELETE THIS LINE --->
<!--- METADATA ANCHOR: MA22019-CW1-2026 --->

Please read the `coursework_01_instructions.pdf` for the full context, data dictionary, and detailed task requirements before starting.



```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(tidytext)
library(knitr)
```

## Part A: Data Wrangling & EDA 

### Task 1: Data Joining
<!--- DO NOT DELETE THIS LINE - TASK 1 ANCHOR --->

```{r}
#| label: task-1-code
# Write your Task 1 code here:

history <- read_csv("data/history.csv")
titles <- read_csv("data/titles.csv")
users <- read_csv("data/users.csv")

history_clean <- history %>% unique ()
titles_clean <- titles %>% unique()
users_clean <- users %>% unique()

cinemetrics <- history_clean %>% 
                  left_join (users_clean, by = 'user_id') %>% 
                    left_join(titles_clean , by = 'movie_id')

nrow(history)
nrow(history_clean)
nrow(titles)
nrow(titles_clean)
nrow(cinemetrics)

```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It was necessary to remove duplicates from history to avoid counting the same viewing record more than once, and duplicates had to be removed from the titles data frame because otherwise repeated movie_id rows would create duplicate matches in the join and artificially increase the final row count.


### Task 2: Missing Values Handling
<!--- DO NOT DELETE THIS LINE - TASK 2 ANCHOR --->

```{r}
#| label: task-2-code
# Write your Task 2 code here:

missing_summary <- tibble (
  variable = colnames ( cinemetrics) , 
  missing_values = colSums(is.na(cinemetrics))
) %>%  arrange(desc(missing_values))

missing_summary %>% knitr::kable(caption = "Missing values per variable")

user_rating_median <- median(cinemetrics$user_rating_100 , na.rm = TRUE)
user_rating_median

```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

It is important to use explicit missing-value functions because they directly identify and count NA values, whereas generic overview functions may not clearly report missingness, which can lead to misinterpretation of the dataset and misleading summary statistics such as the mean, median, or mode; this issue is especially common for character variables, where missing values are often not clearly flagged.This happens because those columns are treated as text rather than numerical variables, so the output may only preview values instead of explicitly counting NAs.


### Task 3: Genre Grouping
<!--- DO NOT DELETE THIS LINE - TASK 3 ANCHOR --->

```{r}
#| label: task-3-code
# Write your Task 3 code here:

cinemetrics |>
  count(genre_tags, sort = TRUE)

cinemetrics <- cinemetrics %>% 
  mutate ( primary_genre = genre_tags %>% 
             as_factor() %>% 
             fct_collapse(
               'Animation' = c("animated", "animation", "animation, kids"),
               'Comedy' = c("comedy", "comedy, family"),
              'Sci-Fi' = c("Science fiction", "sci-fi", "sci-fi, fantasy"),
              'Drama' = c("Drama", "drama, emotional", "drama, critically-acclaimed"),
              'Horror' = c("horror", "scary", "terror"),
              'Romance' = c("Romance", "rom-com", "romantic", "romantic comedy"),
              'Action' = c("action", "action, high-octane"),
              'Documentary' = c("Documentary", "documentary, real-world")
              )
  )

cinemetrics <-  cinemetrics %>%
  mutate( primary_genre = fct_lump(primary_genre , n = 5 , other_level = 'other'))

levels(cinemetrics$primary_genre)
cinemetrics %>% count(primary_genre, sort = TRUE)



```

```{r}
#| label: task-3-verification
# Write your verification code here:

rating_verification <- cinemetrics %>% 
  filter ( genre_tags == 'Romance' | genre_tags =='romantic comedy') %>% 
  group_by( genre_tags) %>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE),
              n = n())

rating_verification %>%
  knitr::kable(caption = "Comparison of raw Romance and romantic comedy genre tags")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

Collapsing the romantic comedy subgroup into the broader Romance category masks its lower average rating, because its mean score of 66.35 across 145 observations is diluted within the much larger Romance group, which has a higher average rating of 69.83 across 991 observations. This is called aggregation distortion.


### Task 4: Genre Ratings Table
<!--- DO NOT DELETE THIS LINE - TASK 4 ANCHOR --->

```{r}
#| label: task-4-code
# Write your Task 4 code here:

genre_ratings <- cinemetrics %>% 
  group_by(primary_genre)%>% 
  summarize ( avg_user_rating_100 = mean(user_rating_100,na.rm = TRUE)) %>%   arrange(desc (avg_user_rating_100))

genre_ratings %>% knitr::kable ( caption = "Average User rating for Primary genre")


```

```{r}
#| label: task-4-verification
# Write your verification code here:

horror_budget <- cinemetrics %>% 
    filter ( genre_tags == "horror" | genre_tags == "scary" |genre_tags== "terror" ) %>% 
    summarise( genre = 'Horror' , avg_budget = mean(budget_usd, na.rm= TRUE)
                 )

action_budget <- cinemetrics %>% 
  filter ( genre_tags == "action" | genre_tags == "action, high-octane" ) %>% 
  summarise( genre = "Action" , avg_budget = mean(budget_usd , na.rm = TRUE ))

SciFi_budget <- cinemetrics %>% 
  filter ( primary_genre == "Sci-Fi") %>% 
  summarise( genre = 'Sci-Fi' , avg_budget = mean (budget_usd, na.rm = TRUE) )

avg_budget_usd = bind_rows (horror_budget,action_budget,SciFi_budget ) %>% 
  arrange ( desc(avg_budget))

avg_budget_usd %>%  knitr::kable (caption = " Average production budget (USD)")


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

The empirical results show that a higher production budget does not guarantee a higher average user rating, because Sci-Fi has the highest average budget of 139588009 USD yet does not have the highest average rating (69.69) in the genre ratings table, where Other performs better(74.34) . However, Action and Horror are absorbed into Other, so their separate average ratings cannot be directly observed from the collapsed table. This demonstrates how a massive generalised category can hide subgroup nuances. The poor performance of a small niche film group, may be concealed when it is absorbed into a larger category whose overall mean is driven by other films.


### Task 5: Ratings and Engagement Scores
<!--- DO NOT DELETE THIS LINE - TASK 5 ANCHOR --->

```{r}
#| label: task-5-code
# Write your Task 5 code here:


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your interpretation of the boxplots' visual footprints.*

```{r}
#| label: task-5-verification
# Write your verification code here:


```

#### Empirical Verification

*Replace this text with your empirical finding regarding the highest and lowest IQRs.*


### Task 6: Engagement Score and Subscription Type
<!--- DO NOT DELETE THIS LINE - TASK 6 ANCHOR --->

```{r}
#| label: task-6-code
# Write your Task 6 code here:


```

```{r}
#| label: task-6-verification
# Write your verification code here:


```

#### Empirical Verification

*Replace this text with your empirical finding regarding the distribution of users across age and subscription tiers.*

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your explanation of the mathematical illusion.*


---

## Part B: Text Data Analysis

### Task 7: Tokenization & Word Frequency
<!--- DO NOT DELETE THIS LINE - TASK 7 ANCHOR --->

```{r}
#| label: task-7-code
# Write your Task 7 code here:


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your explanation of why the top two words are unhelpful for analysis.*


### Task 8: Sentiment Analysis
<!--- DO NOT DELETE THIS LINE - TASK 8 ANCHOR --->

```{r}
#| label: task-8-code
# Write your Task 8 code here:


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your explanation of the scatterplot outlier.*


### Task 9: Advanced Text Mining (TF-IDF)
<!--- DO NOT DELETE THIS LINE - TASK 9 ANCHOR --->

```{r}
#| label: task-9-code
# Write your Task 9 code here: 


```

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

1. *Replace this text with the TF-IDF score for the word "film".*
2. *Replace this text with your mathematical explanation using the IDF equation from the lecture.*
3. *Replace this text with your comparison of TF-IDF vs frequency.*



### Task 10: Executive Recommendation
<!--- DO NOT DELETE THIS LINE - TASK 10 ANCHOR --->

<!--- WRITE YOUR WRITTEN ANSWER BELOW THIS LINE --->

#### Written Interpretation

*Replace this text with your one-paragraph executive recommendation (max 150 words), citing at least three specific findings from earlier tasks and acknowledging one limitation.*


---
<!--- WORD COUNT ANCHOR --->
```{r}
#| label: word-count
#| echo: false
#| message: false
#| warning: false

count_prose_words <- function(file_path) {
  # Read the raw lines of the file
  lines <- readLines(file_path, warn = FALSE)
  
  # 1. Strip YAML Header (Lines between the first two '---')
  yaml_bounds <- grep("^---$", lines)
  if (length(yaml_bounds) >= 2) {
    lines <- c(lines[1:(yaml_bounds[1]-1)], lines[(yaml_bounds[2]+1):length(lines)])
  }
  
  # 2. Strip Code Chunks (Lines between pairs of '```')
  chunk_bounds <- grep("^```", lines)
  if (length(chunk_bounds) %% 2 == 0 && length(chunk_bounds) > 0) {
    # Create a vector of line indices to remove
    lines_to_remove <- integer(0)
    for (i in seq(1, length(chunk_bounds), by = 2)) {
      lines_to_remove <- c(lines_to_remove, chunk_bounds[i]:chunk_bounds[i+1])
    }
    lines <- lines[-lines_to_remove]
  }
  
  # 3. Strip Markdown Headers (Lines starting with #)
  lines <- lines[!grepl("^#+", trimws(lines))]
  
  # 4. Strip HTML Comments (Lines starting with <!--)
  lines <- lines[!grepl("^<!--", trimws(lines))]
  
  # 5. Combine what's left into a single string
  text_only <- paste(lines, collapse = " ")
  
  # 6. Remove common markdown characters that might be counted as words
  text_only <- gsub("[*>`-]", " ", text_only)
  
  # 7. Count the remaining words
  words <- unlist(strsplit(trimws(text_only), "\\s+"))
  words <- words[words != ""]
  
  return(length(words))
}

# Run the function on the current file
true_word_count <- count_prose_words(knitr::current_input())

# Print the final result formatted nicely
cat(sprintf("**Prose Word Count:** %d words", true_word_count))
```
